HGP analysis: fit model

### 2019-03-26 legacy.tex

To build a test statistic to represent the association across all populations simultaneously, we performed a simple logistic regression predicting genotype based on $Q$ with population as a covariate to account for population structure.
This method identifies a total of 11,416 variants associated to $Q$ (Figures \ref{NRS_Manhattan},\ref{NRI_Manhattan}, \ref{RS_Manhattan}, and \ref{RI_Manhattan}): 
To account for the large number of tests, we used a two-stage Benjamini \& Hochberg step-up FDR-controlling procedure to adjust the p-values using a nominal Type-I error rate $\alpha = 0.01$ \citep{Benjamini2006}. We tested SNPs, INDELs and repetitive regions separately as they may have different error rates (Table \ref{sigTable}).

\begin{table}[h]
\centering
\begin{tabular}{l  r r}
                      & {Repeat}  & {Non-Repeat}       \\ \hline
{SNP}  & 2,007 & 8,009 \\  
{INDEL} & 330 & 1,070 \\ \hline
\end{tabular}
\caption{Number of statistically significant variants per category. Variants that are flagged by the 1kGP nested repeat mask file were analyzed separately. SNPs and INDELs were also analyzed separately. The total number of 11,416 are statistically significantly associated to $Q$.}
\label{sigTable}
\end{table}

>>>>

To build a test statistic to represent the association across all populations simultaneously, we performed a simple logistic regression predicting genotype based on $Q$ with population as a covariate to account for population structure.

This method identifies a total of 24,390 variants associated to $Q$ (Figures \ref{NRS_Manhattan},\ref{NRI_Manhattan}, \ref{RS_Manhattan}, and \ref{RI_Manhattan}). 
To account for the large number of tests, we used a two-stage Benjamini \& Hochberg step-up FDR-controlling procedure to adjust the p-values using a nominal Type-I error rate $\alpha = 0.01$ \citep{Benjamini2006}. We tested SNPs, INDELs and repetitive regions separately as they may have different error rates (Table \ref{sigTable}).

\begin{table}[h]
\centering
\begin{tabular}{l  r r}
                      & {Repeat}  & {Non-Repeat}       \\ \hline
{SNP}  & 5,186 & 17,124 \\  
{INDEL} & 509 & 1,571 \\ \hline
\end{tabular}
\caption{Number of statistically significant variants per category. Variants that are flagged by the 1kGP nested repeat mask file were analyzed separately. SNPs and INDELs were also analyzed separately. The total number of 24,390 are statistically significantly associated to $Q$.}
\label{sigTable}
\end{table}

### 2019-03-26 rick/final/3_fit_model_compareModels.Rmd

#chr <- args[3]
chr <- '22_10000'

@@

```{r}
#this runs the logistic regression
lreg_dich <- function(y, X){

  b <- glm2(cbind(y, 2-y) ~ -1 + X, family="binomial")$coef

}

#modified from gcat
fit_glm_dich <- function(..., LF, trait){
  snp <- c(...)
  snp <- as.numeric((snp==1) | (snp==2))
  ind <- !is.na(snp) #logical index vector
  snp_no_na <- snp[ind]
  LF_no_na <- LF[ind,]
  p0 <- rep(NA, length(snp))
  p1 <- rep(NA, length(snp))
  trait_no_na <- trait[ind]
  b0 <- lreg_dich(snp_no_na, LF_no_na) #coefficients from logreg 
  b1 <- lreg_dich(snp_no_na, cbind(LF_no_na, trait_no_na))
  est0 <- .Call("mv", LF_no_na, b0)
  est1 <- .Call("mv", cbind(LF_no_na, trait_no_na), b1)
  p0[ind] <- exp(est0)/(1+exp(est0))
  p1[ind] <- exp(est1)/(1+exp(est1))

  stat <-
    dataset %>%
    summarize( #cross entropy (log likelihood)
      dev0 = -2 * sum(snp * log(p0) +
                        (1 - snp) * log(1 - p0),
                      na.rm = TRUE),
      dev1 = -2 * sum(snp * log(p1) +
                        (1 - snp) * log(1 - p1),
                      na.rm = TRUE),
      devdiff = dev0 - dev1,
      dev_logp = -log10(pchisq(devdiff,
                               1,
                               lower.tail = FALSE))
    ) %>% 
    mutate(num_nas = sum(!ind),
	   num_zeros= sum(snp_no_na ==0),
           b=b1[length(b1)])

define functions
```{r}

#this runs the logistic regression
lreg <- function(y, X){
  X <- rbind(X, X)
  y1 <- as.numeric((y==1) | (y==2))
  y2 <- as.numeric(y==2)
  y <- c(y1,y2)
  b <- glm2(cbind(y, 2-y) ~ -1 + X, family="binomial")$coef

@@

data_fit_dich <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm_dich,
	LF= LF, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()    
```

@@

saveRDS(data_fit, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, ".rds")))

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)

saveRDS(data_fit_dich, 
	file.path("./data/preprocessed/subsets",
	paste0("dich_fits_chrom_", chr, ".rds")))

write.table(data_fit_dich[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("dich_fits_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)
```
#PCA
```{r, warning=FALSE}
#load PCA
library(data.table)
samples = fread('/Users/luke/Documents/PCAperPop/Name_Pop_Qual_PC1_PC2_PC3_PC4_PC5.txt')
PC <- as.matrix(samples[,c('PC1','PC2','PC3','PC4','PC5')])
PC <- cbind(PC, 1)

@@

saveRDS(data_fit_PCA, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, ".rds")))

write.table(data_fit_PCA[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fitsPCA_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)

saveRDS(data_fit_PCA_dich, 
	file.path("./data/preprocessed/subsets",
	paste0("dich_fits_chrom_", chr, ".rds")))

write.table(data_fit_PCA_dich[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("dich_fitsPCA_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)
```
#Population
```{r, warning=FALSE}

@@

data_fit_Pop_dich <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm_dich,
	LF= testModel, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()   
```
```{r}

saveRDS(data_fit_Pop, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, ".rds")))

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fitsPOP_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)

saveRDS(data_fit_Pop_dich, 
	file.path("./data/preprocessed/subsets",
	paste0("dich_fits_chrom_", chr, ".rds")))

write.table(data_fit_dich[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("dich_fitsPOP_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)
```

>>>>

#chr <- args[3]
chr <- '22'

@@

define functions
```{r}
#this runs the logistic regression
lreg <- function(y, X){
  X <- rbind(X, X)
  y1 <- as.numeric((y==1) | (y==2))
  y2 <- as.numeric(y==2)
  y <- c(y1,y2)
  b <- glm2(cbind(y, 2-y) ~ -1 + X, family="binomial")$coef

@@

@@

saveRDS(data_fit, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_LFA.rds")))

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_LFA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit[,2]<-NULL
```
#PCA
```{r, warning=FALSE}
#load PCA
library(data.table)
samples = fread('/Users/luke/Documents/PCAperPop/Name_Pop_Qual_PC1_PC2_PC3_PC4_PC5.txt')
PC <- as.matrix(samples[,c('PC1','PC2','PC3','PC4','PC5')])
PC <- cbind(PC, 1)

@@

saveRDS(data_fit_PCA, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_PCA.rds")))

write.table(data_fit_PCA[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_PCA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit_PCA[,2]<-NULL
```
#Population
```{r, warning=FALSE}

@@

```
```{r}

saveRDS(data_fit_Pop, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_Pop.rds")))

write.table(data_fit_Pop[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_Pop.txt")),
              quote = FALSE,
              row.names = FALSE)

data_fit_Pop[,2]<-NULL
```

### 2019-03-26 rick/final/Luke_Cluster/3_fit_model.Rmd

#path <- args[1]
path <- '/Users/luke/Documents/QualityPaper/rick/final/'
setwd(path)

#plinkPath <- args[2] 
plinkPath <- '/Users/luke/genomes/genomes/hg19/plink/'

#chr <- args[3]
chr <- '22'

@@

```{r}
library(tidyverse)
library(lfa)
library(furrr)
library(glm2)
library(tictoc)
library(data.table)
library(cowplot)
options(future.globals.maxSize= 1000000000000)
plan(multicore)
```
load plink format genotype data
```{r}
data_genotype <- read.bed(paste0(plinkPath,'chr',chr))

@@

rownames(data_genotype) <- data_bim$rsID
```
run lfa if needed
```{r lfa}
tic()
#LF <- lfa(data_genotype, 5)
toc()
message("lfa run complete")
```
save lfa if needed
```{r save}
tic()
#saveRDS(LF, "./data/preprocessed/lf5.rds")
toc()
```
load logistic factor analysis results
```{r read lf}
LF <- readRDS( "./data/preprocessed/lf5.rds")
```

define functions
```{r}
#this runs the logistic regression
lreg <- function(y, X){
  X <- rbind(X, X)
  y1 <- as.numeric((y==1) | (y==2))
  y2 <- as.numeric(y==2)
  y <- c(y1,y2)
  b <- glm2(cbind(y, 2-y) ~ -1 + X, family="binomial")$coef

@@

}
#modified from gcat
fit_glm <- function(..., LF, trait){
  snp <- c(...)
  ind <- !is.na(snp) #logical index vector
  snp_no_na <- snp[ind]
  LF_no_na <- LF[ind,]
  p0 <- rep(NA, length(snp))
  p1 <- rep(NA, length(snp))
  trait_no_na <- trait[ind]

  b0 <- lreg(snp_no_na, LF_no_na) #coefficients from logreg 
  b1 <- lreg(snp_no_na, cbind(LF_no_na, trait_no_na))
  est0 <- .Call("mv", LF_no_na, b0)
  est1 <- .Call("mv", cbind(LF_no_na, trait_no_na), b1)
  p0[ind] <- exp(est0)/(1+exp(est0))
  p1[ind] <- exp(est1)/(1+exp(est1))

@@

```{r}
head(avg_qual_mean)
```
#LFA
```{r, warning=FALSE}
tic("fitting model")
labels <- rownames(data_genotype)
data_fit <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm,
	LF=LF, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)

@@

saveRDS(data_fit, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_LFA.rds")))

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_LFA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit[,2]<-NULL
```
#PCA
```{r, warning=FALSE}
#load PCA
library(data.table)
samples = fread('/Users/luke/Documents/PCAperPop/Name_Pop_Qual_PC1_PC2_PC3_PC4_PC5.txt')
PC <- as.matrix(samples[,c('PC1','PC2','PC3','PC4','PC5')])
PC <- cbind(PC, 1)

tic("fitting model")
labels <- rownames(data_genotype)
data_fit_PCA <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm,
	LF= PC, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)

data_fit_PCA_dich <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm_dich,
	LF= PC, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()    
```

saveRDS(data_fit_PCA, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_PCA.rds")))

write.table(data_fit_PCA[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_PCA.txt")),
              quote = FALSE,
              row.names = FALSE)
data_fit_PCA[,2]<-NULL
```
#Population
```{r, warning=FALSE}

testModel <- model.matrix(data = bas_dt, object = ~Population)

tic("fitting model")
labels <- rownames(data_genotype)
data_fit_Pop <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm,
	LF= testModel, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()

```
```{r}

saveRDS(data_fit_Pop, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, "_Pop.rds")))

write.table(data_fit_Pop[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, "_Pop.txt")),
              quote = FALSE,
              row.names = FALSE)

data_fit_Pop[,2]<-NULL
```

@@

```{r,fig.height = 10, fig.width = 10}

plt <- merge(data_fit[-2], 
             data_fit_PCA[-2], 
             by=c('rsID'), 
             suffixes = c('LFA','PCA'))
plt <- merge(plt, 
             data_fit_Pop[-2], 
             by=c('rsID'), 
             suffixes = c('','Pop'))

beta1<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
beta2<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
beta3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
betas <- plot_grid(beta1,beta2,beta3,nrow=1)

af1<-ggplot(plt, aes(1-num_zeros/2504, bLFA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='LFA')+ggtitle('Beta')+xlim(c(0,0.3))
af2<-ggplot(plt, aes(1-num_zeros/2504, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='PCA')+ggtitle('Beta')+xlim(c(0,0.3))
af3<-ggplot(plt, aes(1-num_zeros/2504, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='AF', y='Pop')+ggtitle('Beta')+xlim(c(0,0.3))
af <- plot_grid(af1,af2,af3,nrow=1)
af
ggsave('./Beta_AF.jpg', height=4, width=8)

plt <- plt[which(plt$num_zeros < 2501),]

beta1<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
beta2<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
beta3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
betas2 <- plot_grid(beta1,beta2,beta3,nrow=1)

af1<-ggplot(plt, aes(1-num_zeros/2504, bLFA))+geom_point(shape=1)+labs(x='AF', y='LFA')+ggtitle('Beta')+xlim(c(0,0.3))
af2<-ggplot(plt, aes(1-num_zeros/2504, bPCA))+geom_point(shape=1)+labs(x='AF', y='PCA')+ggtitle('Beta')+xlim(c(0,0.3))
af3<-ggplot(plt, aes(1-num_zeros/2504, b))+geom_point(shape=1)+labs(x='AF', y='Pop')+ggtitle('Beta')+xlim(c(0,0.3))
af2 <- plot_grid(af1,af2,af3,nrow=1)
af2
ggsave('./Beta_AF2.jpg', height=4, width=8)

makePlot <- function(x,y,xlab,ylab,title) {
p <- ggplot(plt, aes(x, y)) + 
  geom_abline(color='grey') +
  geom_point(shape=1) +
  labs(x=xlab, y=ylab) +
  ggtitle(title)

return(p)
}

p1 <- makePlot(plt$dev0LFA, 
               plt$dev0PCA, 
               'Null Deviance LFA', 
               'Null Deviance PCA', 
               'Null Model')
p2 <- makePlot(plt$devdiffLFA, 
               plt$devdiffPCA, 
               'Deviance LFA', 
               'Deviance PCA', 
               'Deviance')

p2.2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))

p3<-ggplot(plt, aes(bLFA, bPCA, color=1-num_zeros/2504))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0LFA, dev0))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, b, color=1-num_zeros/2504))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,50))+ylim(c(0,50))
p3<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels.jpg', height=10, width = 10)

```{r,fig.height = 10, fig.width = 10}

plt <- merge(data_fit_dich[-2], data_fit_PCA_dich[-2], by=c('rsID'), suffixes = c('LFA','PCA'))
plt <- merge(plt, data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('','Pop'))

p1<-ggplot(plt, aes(dev0LFA, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='PCA')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0LFA, dev0))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiffLFA, devdiff))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(plt, aes(bLFA, b))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA', y='Pop')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(plt, aes(dev0, dev0PCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Null Model')
p2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance')
p2.2<-ggplot(plt, aes(devdiff, devdiffPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Deviance Zoom')+xlim(c(0,50))+ylim(c(0,50))
p3<-ggplot(plt, aes(b, bPCA))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Pop', y='PCA')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels_dich.jpg', height=10, width = 10)

```{r,fig.height = 10, fig.width = 10}

LFAp <- merge(data_fit[-2], data_fit_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))
PCAp <- merge(data_fit_PCA[-2], data_fit_PCA_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))
Popp <- merge(data_fit_Pop[-2], data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('Poly','Dich'))

p1<-ggplot(LFAp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model LFA')
p2<-ggplot(LFAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(LFAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(LFAp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(PCAp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model PCA')
p2<-ggplot(PCAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(PCAp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(PCAp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp2 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

p1<-ggplot(Popp, aes(dev0Poly, dev0Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Null Model Pop')
p2<-ggplot(Popp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance')
p2.2<-ggplot(Popp, aes(devdiffPoly, devdiffDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))
p3<-ggplot(Popp, aes(bPoly, bDich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='Poly', y='Dich')+ggtitle('Beta')
pp3 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

plot_grid(pp1,pp2,pp3,ncol=3)
ggsave('./CompareModels_VSdich.jpg', height=10, width = 10)

```{r,fig.height = 10, fig.width = 6}
#plot diff
plt <- merge(data_fit[-2], data_fit_Pop_dich[-2], by=c('rsID'), suffixes = c('LFA_Full','Pop_Dich'))

p1<-ggplot(plt, aes(dev0LFA_Full, dev0Pop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA Full', y='Pop Dich')+ggtitle('Null Model')+coord_flip()
p2<-ggplot(plt, aes(devdiffLFA_Full, devdiffPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Deviance')+coord_flip()
p2.2<-ggplot(plt, aes(devdiffLFA_Full, devdiffPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Deviance Zoom')+xlim(c(0,30))+ylim(c(0,30))+coord_flip()
p3<-ggplot(plt, aes(bLFA_Full, bPop_Dich))+geom_abline(color='grey')+geom_point(shape=1)+labs(x='LFA_Full', y='Pop_Dich')+ggtitle('Beta')+coord_flip()
pp1 <- plot_grid(p1,p2,p2.2,p3,ncol=1)

pp1
ggsave('./CompareModels_FullLFA_Popdich.jpg', height=10, width = 6)
```

>>>>

path <- args[1]
#path <- '/Users/luke/Documents/QualityPaper/rick/final/'
setwd(path)

plinkPath <- args[2] 
#plinkPath <- '/Users/luke/genomes/genomes/hg19/plink/'

chr <- args[3]
#chr <- '22_10000'

@@

```{r}
library(tidyverse)
library(lfa)
library(furrr)
library(glm2)
library(tictoc)
library(data.table)
options(future.globals.maxSize= 1000000000000)
plan(multicore)
```
load plink format genotype data
```{r}
data_genotype <- read.bed(paste0(plinkPath,'chr',chr))

@@

rownames(data_genotype) <- data_bim$rsID
```
run lfa if needed
```{r lfa}
tic()
#LF <- lfa(data_genotype, 5)
toc()
message("lfa run complete")
```
save lfa if needed
```{r save}
tic()
#saveRDS(LF, "./data/preprocessed/lf5.rds")
toc()
```
load logistic factor analysis results
```{r read lf}
LF <- readRDS( "./data/preprocessed/lf5.rds")
```
define functions
```{r}
#this runs the logistic regression
lreg <- function(y, X){
  X <- rbind(X, X)
  y1 <- as.numeric((y==1) | (y==2))
  y2 <- as.numeric(y==2)
  y <- c(y1,y2)
  b <- glm2(cbind(y, 2-y) ~ -1 + X, family="binomial")$coef

@@

}
#modified from gcat
fit_glm <- function(..., LF, trait){
  snp <- c(...)
  ind <- !is.na(snp) #logical index vector
  snp_no_na <- snp[ind]
  LF_no_na <- LF[ind,]
  p0 <- rep(NA, length(snp))
  p1 <- rep(NA, length(snp))
  trait_no_na <- trait[ind]
  b0 <- lreg(snp_no_na, LF_no_na) #coefficients from logreg 
  b1 <- lreg(snp_no_na, cbind(LF_no_na, trait_no_na))
  est0 <- .Call("mv", LF_no_na, b0)
  est1 <- .Call("mv", cbind(LF_no_na, trait_no_na), b1)
  p0[ind] <- exp(est0)/(1+exp(est0))
  p1[ind] <- exp(est1)/(1+exp(est1))

@@

```{r}
head(avg_qual_mean)
```

```{r}
tic("fitting model")
labels <- rownames(data_genotype)
data_fit <- 
	data_genotype %>%
	as_tibble() %>%
	transmute(
	fit = future_pmap(., fit_glm,
	LF=LF, 
	trait=avg_qual_mean),
		rsID=labels)  %>%
	unnest(fit) %>%
	mutate(l= rep(c("data","stat"), n()/2))%>%
	spread(l, fit)%>%
	unnest(stat)
   toc()

@@

saveRDS(data_fit, 
	file.path("./data/preprocessed/subsets",
	paste0("fits_chrom_", chr, ".rds")))

write.table(data_fit[-2], 
              file.path("./data/preprocessed/subsets", 
                        paste0("fits_chrom_", chr, ".txt")),
              quote = FALSE,
              row.names = FALSE)
```

@@
